Testing the Relationship between Phenomenological Control related to Illusion Sensitivity

### 2024-06-07 paper/R0/manuscript.qmd
edits
@AnafNeves made some edits in the revision, can you transfer to ms and knit

Interestingly, @lush2022 did test for a relationship between PC and illusion sensitivity using the Müller-Lyer illusion (in which the arrangement of the arrowheads flanking two lines makes them appear as having different lengths), and reported evidence in favour of an absence of correlation between the two measures. This finding was interpreted as indicative of the cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC. The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that would encompass other illusion types (see @tbl-table1).**Our paradigm extends that of Lush’s by presenting this task as a gamified paradigm hopefully motivating participants to perform at their best and to remain engaged throughout**.

@@

| **Question**                                                                                                                                 | **Hypothesis**                                                                                                                            | **Sampling Plan**                                                                                                               | **Analysis Plan**                                                                                                                                                                             | **Rationale for Deciding the Sensitivity of the Test**                                                                                           | **Interpretation Given Different Outcomes**                         | **Theory That Could Be Shown Wrong by the Outcomes**                                                                                                        |
|----------------------------------------------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------------------------------------------------------------------------|--------------------------------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|-------------------------------------------------------------------------------------------------------------------------------|--------------------------------------------------------------------|----------------------------------------------------------------------------------------------------------------------------------------------------|
| Is there a correlation between trait phenomenological control (PC) and visual illusion (VI) sensitivity?                                    | Replicating findings from Lush et al., 2022 paper, we expect evidence of favour of an absence of relationship between VI and PCS         | The goal is to recruit around 500 adult English speakers using Prolific. This sample size is based on the ones used in Lush et al., 2021 and Lush et al., 2022 that we aim at replicate. | Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the BayesFactor::BFCorrelation() function (with the r-scale prior parameter set to 'medium') | Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings. BF10 > 3 will be interpreted as evidence for a relationship between these two measures. | The hypothesis that VI sensitivity is independent from PC.          | The cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC. |

>>>>

Interestingly, @lush2022 did test for a relationship between PC and illusion sensitivity using the Müller-Lyer illusion (in which the arrangement of the arrowheads flanking two lines makes them appear as having different lengths), and reported evidence in favour of an absence of correlation between the two measures. This finding was interpreted as indicative of the cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC. The goal of this study is thus to replicate the results from @lush2022 pointing to an absence of a relationship between phenomenological control and illusion sensitivity, by generalising them to a different illusion task that encompasses other illusion types (see @tbl-table1). **Our paradigm extends that of Lush’s by presenting this task as a gamified paradigm hopefully motivating participants to perform at their best and to remain engaged throughout**.

@@

| **Question**                                                                                                                                 | **Hypothesis**                                                                                                                            | **Sampling Plan**                                                                                                               | **Analysis Plan**                                                                                                                                                                             | **Rationale for Deciding the Sensitivity of the Test**                                                                                           | **Interpretation Given Different Outcomes**                         | **Theory That Could Be Shown Wrong by the Outcomes**                                                                                                        |
|----------------------------------------------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------------------------------------------------------------------------|--------------------------------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|-------------------------------------------------------------------------------------------------------------------------------|--------------------------------------------------------------------|----------------------------------------------------------------------------------------------------------------------------------------------------|
| Is there a correlation between trait phenomenological control (PC) and visual illusion (VI) sensitivity?                                    | Replicating findings from Lush et al., 2022 paper, we expect evidence of favour of an absence of relationship between VI and PCS         | The goal is to recruit around 500 adult English speakers using Prolific. This sample size is based on the ones used in Lush et al., 2021 and Lush et al., 2022 that we aim at replicate. | Bayesian correlations between the PC score and the VI performance for the 3 illusion types (corresponding to the error rate) will be computed using the `BayesFactor::BFCorrelation()` function (with the r-scale prior parameter set to 'medium') | Evidence against a relationship between PC and VI will be found if BF10 <= 1/3, following the Lush et al., 2022 findings. BF10 > 3 will be interpreted as evidence for a relationship between these two measures. | The hypothesis that VI sensitivity is independent from PC.          | The cognitive impenetrability of illusions, implying that the effect is driven by low-level processes and therefore not influenced by top-down mechanisms such as PC. |
